---
title: "Question 5"
output: html_document
date: '2022-03-28'
---

```{r}
#Importing the combined dataset
datacomb=read.csv("../Coursework R/datacomb.csv")
```

```{r}
#Checking for missing values in "datacomb" before checking correlation and finding a model
colSums(is.na(datacomb))
```

```{r}
#Removing the unnecessary column named "X" from "datacomb"
datacomb<-datacomb[,2:30]
#Viewing the top 5 rows of "datacomb"
head(datacomb)
```

```{r}
#Removing missing values only from the "ArrDelay" column only for now since the dependent variable of the model is going to be ArrDelay 
datacomb<-datacomb[!is.na(datacomb$ArrDelay),]
```

```{r}
#Exploring "datacomb" including checking the datatypes of the column variables
summary(datacomb)
```

```{r}
#Taking out the categorical variables out of 'datacomb' to plot the correlation plot
data<-datacomb[,c(1,2,3,4,5,6,7,8,12,13,14,15,16,19,20,21,25,26,27,28,29)]
```

```{r}
#Viewing the head of 'data' to check whether the columns in 'data' are numerical variables
head(data)
```

```{r}
#Checking the correlation between the column variables in "datacomb"
#Importing necessary libraries
library(plotly)
library(ggplot2)
library(corrplot) #Importing the "corrplot" library
corrplot(abs(cor(data)), method="color",
         title="A correlation plot for columns in 'data'",
         tl.col="black" , tl.srt=45 ,
         addCoef.col="black" ,
         number.cex = 0.5 ,
         tl.cex = 0.5 ,
         is.corr= FALSE ,
         col=colorRampPalette(c("khaki","lightblue1","red"))(100))
```

## A multiple linear regression model is constructed

```{r}
#Creating a dataframe named 'data1' to store the columns that are needed for the regression model
data1<-data[,c(2,3,4,6,12,13,21)]
```

```{r}
#Checking for missing values in 'data1' before we create the model
colSums(is.na(data1))
```

###No missing values in 'data1'

```{r}
#Importing the "catools" library and split "data1" into training and testing datasets
library(caTools)
set.seed(1)
#Creating a vector named 'split' containing 0s and 1s to make 70% training and 30% testing sets
split<- sample(c(rep(0, 0.7 * nrow(data1)), rep(1, 0.3 * nrow(data1))))
```

```{r}
#Getting the training and testing datasets using the 'split' vector
train <- data1[split == 0, ] #Creating the training set
test <- data1[split== 1, ]   #Creating the testing set
```

```{r}
#Viewing the first few rows of "train" and "test"
head(train)
head(test)
```

```{r}
#Seperating the explanatory(independent) and target(dependent) variables from the training and testing datasets
x_train<-train[,c(1,2,3,4,6,7)]
y_train<-train[,5]
x_test<-test[,c(1,2,3,4,6,7)]
y_test<-test[,5]
```

```{r}
#Scaling the explanatory variables in testing and training sets to avoid multicollinearity
x_train<-scale(x_train)
x_test<-scale(x_test)
```
### y_test and y_train are not scaled since it is the target variable "ArrDelay"

```{r}
#Converting "train" and "test" to dataframes
library(dplyr)
x_train<-as.data.frame(x_train)
x_test<-as.data.frame(x_test)
```

```{r}
#Combining y_train with the scaled x_train to form the scaled training set
scaled_train<-cbind(x_train,y_train)
#Combining y_test with the scaled x_test to form the scaled testing test
scaled_test<-cbind(x_test,y_test)
```

```{r}
#Viewing the first few rows of the scaled training and testing datasets
head(scaled_test)
head(scaled_train)
```

```{r}
#Renaming the column "y_train" in "scaled_train" and the column "y_test" in "scaled_test" to "ArrDelay"
colnames(scaled_train)<-c("Month","DayofMonth","DayOfWeek","CRSDepTime","DepDelay","LateAircraftDelay","ArrDelay")
colnames(scaled_test)<-c("Month","DayofMonth","DayOfWeek","CRSDepTime","DepDelay","LateAircraftDelay","ArrDelay")
```

```{r}
#Building the multiple linear regression model
Multilinearmodel<-lm(formula=ArrDelay~Month+DayofMonth+DayOfWeek+CRSDepTime+DepDelay+LateAircraftDelay, data = scaled_train)
```

```{r}
#Evaluation of the model
summary(Multilinearmodel)
```

#Finding the Mean Squared Error for the model

```{r}
#Finding the predicted Arrivaldelay using the testing dataset "x_test"
y_pred<-predict(Multilinearmodel,newdata = x_test)
```

```{r}
#Viewing "y_pred"
head(y_pred)
```
```{r}
#Comparing the predicted y values from the model with the y_test data
head(y_test)
```
###Seems to have variations in the y values but the 6th observation is really close

```{r}
#Calculating the mean squared error
MSE<-mean((scaled_test$ArrDelay-y_pred)**2)
RMSE<-sqrt(MSE)
```

```{r}
paste("The mean squared error for this model is", MSE)
paste("The root mean squared error for this model is", RMSE)
```


#Plotting a histogram on residuals to check whether the errors are normally distributed

```{r}
#Creating a vector named 'res' to include the residuals of the model
res<-residuals(Multilinearmodel)
#Converting the residual vector to a dataframe
res<- as.data.frame(res)
```

```{r}
#Plotting the histogram 
ggplot(res,aes(res)) + geom_histogram(fill='blue',alpha=0.5,bins=100) + ggtitle("Histogram of the residuals") + xlim(c(-50,50))
```
