---
title: "Question 2"
output: html_document
date: '2022-03-27'
---

#Importing the plane_data and datacomb datasets
```{r}
datacomb<-read.csv("../Coursework R/datacomb.csv")
plane_data<-read.csv("../Coursework R/planedata.csv")
```

#Renaming tailnum in plane_data dataset to make it common to both datacomb and plane_data
```{r}
colnames(plane_data)[2]<-"TailNum"
```

#Removing missing values from DepDelay and ArrDelay
```{r}
datacomb<-datacomb[!is.na(datacomb$DepDelay), ]
datacomb<-datacomb[!is.na(datacomb$ArrDelay), ]
```

#Creating a dataframe named "dataset" with only the required columns for this question
```{r}
dataset=datacomb[,c(2,12,16,17)]
head(dataset)
```

```{r}
#merging both plane_data and datacomb on TailNum
finalcomb<-merge (dataset,plane_data, by="TailNum")
```
```{r}
#Checking the top 5 rows of the dataframe "finalcomb"
head(finalcomb)
```
```{r}
#Removing the unnecessary column named "X" from "finalcomb"
finalcomb=finalcomb[,c(1,2,3,4,6,7,8,9,10,11,12,13)]
head(finalcomb)
```

#Making a table by the aggregate function with average of DepDelay,average of ArrDelay and Year
```{r}
AvgArrDelay_year<-aggregate(finalcomb$ArrDelay , by=list(finalcomb$year), FUN=mean) 
AvgDepDelay_year<-aggregate(finalcomb$DepDelay , by=list(finalcomb$year), FUN=mean) 
#Column binding AvgArrDelay and AvgDepDelay
Avgdelays=cbind(AvgArrDelay_year,AvgDepDelay_year) 
#Removing the unnecessary columns
Avgdelays<-Avgdelays[,c(1,2,4)] 
```
```{r}
#Column names are changed appropriately
colnames(Avgdelays)<-c('Manufacture_year','AvgArrDelay','AvgDepDelay') 
#Average of delays is calculated once again - the same method as in Q1
Avgdelays['AvgOfDelays'] <- (Avgdelays$AvgArrDelay+Avgdelays$AvgDepDelay)/2
Avgdelays
```
```{r}
#It is seen when we go through "Avgdelays" table above that there are 2 out of range data(meaningless)  under Manufacture_year column which are'0000' and 'None' hence they are removed.
Avgdelays<-Avgdelays[2:49,]
#The manufacture year column is changed into integer datatype to do a line plot and to find the correlation between plane manufacture year and Avgofdelays
Avgdelays$Manufacture_year<-as.integer(Avgdelays$Manufacture_year)
```

#Plotting a line plot to visualize
```{r}
#importing library "ggplot2"
library(ggplot2)
#Line plot
ggplot()+ geom_line(data=Avgdelays, aes(x=Manufacture_year,y=AvgOfDelays, colour="AvgofDelays"))+  geom_point(data=Avgdelays, aes(x=Manufacture_year,y=AvgOfDelays, colour="AvgofDelays")) + xlab("Plane Manufacture Year")+ ylab("Average Delay/mins")+
  ggtitle("Line graph representing Average delays per plane manufacture year")
```

```{r}
#Correlation (spearman) is too found to check whether there is an association between Plane Manufacture Year and the average delay.
cor(Avgdelays$Manufacture_year, Avgdelays$AvgOfDelays,
    method = "pearson")
```
# The correlation between Plane manufacture year and Avgofdelays is -0.07109128

