---
title: "Question1"
output: html_document
date: '2022-03-27'
---

#Importing the cleaned combined dataset "datacomb"
```{r}
datacomb<-read.csv("../Coursework R/datacomb.csv")
```

#removing missing values in DepDelay and ArrDelay since we need these columns for this question
```{r}
data <- datacomb[!is.na(datacomb$DepDelay), ]

data <- data[!is.na(data$ArrDelay), ]
```

## Best DayOfweek to minimise delays

```{r}
#grouping Dayofweek by averagearrivaldelays and average departure delays
AvgArrDelay<-aggregate(data$ArrDelay , by=list(data$DayOfWeek), FUN=mean) 
AvgDepDelay<-aggregate(data$DepDelay , by=list(data$DayOfWeek), FUN=mean) 
```

```{r}
#Column binding both AvgArrDelay and AvgDepDelay variables 
Avgdelaysperweekday=cbind(AvgArrDelay,AvgDepDelay) 
#Removing the common Dayofweek column in the binded variable
Avgdelaysperweekday<-Avgdelaysperweekday[,c(1,2,4)] 
#Renaming the  column names appropriately
colnames(Avgdelaysperweekday)<-c('DayOfWeek','AvgArrDelay','AvgDepDelay') 
#Finding the average of both averages of AvgArrDelay and AvgDepDelay
AvgofDelays <- (Avgdelaysperweekday$AvgArrDelay+Avgdelaysperweekday$AvgDepDelay)/2
#Column binding AvgOfDelays and Avgdelaysperweekday
Avgdelaysperweekday<-cbind(Avgdelaysperweekday,AvgofDelays)
#Viewing "Avgdelaysperweekday" table
show(Avgdelaysperweekday)
```

#Plotting a lineplot to visualise

```{r}
#importing the library "ggplot2" needed for data visualisations
library(ggplot2)
#Plotting a lineplot to find the best dayofweek to minimise delays
ggplot()+ geom_line(data=Avgdelaysperweekday, aes(x=DayOfWeek,y=AvgArrDelay, colour="AvgArrDelay"))+ geom_line(data=Avgdelaysperweekday, aes(x=DayOfWeek,y=AvgDepDelay, colour="AvgDepDelay")) + geom_line(data=Avgdelaysperweekday, aes(x=DayOfWeek,y=AvgofDelays, colour="AvgofDelays")) +
geom_point(data=Avgdelaysperweekday, aes(x=DayOfWeek,y=AvgArrDelay, colour="AvgArrDelay")) + geom_point(data=Avgdelaysperweekday, aes(x=DayOfWeek,y=AvgDepDelay, colour="AvgDepDelay")) + geom_point(data=Avgdelaysperweekday, aes(x=DayOfWeek,y=AvgofDelays, colour="AvgofDelays")) + xlab("DayOfWeek")+ ylab("Average Delay/mins")+ scale_x_discrete(limits=c("Mon","Tue","Wed","Thurs","Fri","Sat","Sun"))+
ggtitle("Line graph representing AverageDelays per weekday")
```
### Saturday is the best day of week to minimse delays

##Best Month to minimise delays

```{r}
#Grouping Months by Average arrival and average departure delays
AvgArrDelay_month<-aggregate(data$ArrDelay , by=list(data$Month), FUN=mean) 
AvgDepDelay_month<-aggregate(data$DepDelay , by=list(data$Month), FUN=mean)
```

```{r}
#Column binding the right above 2 dataframes 
Avgdelayspermonth=cbind(AvgArrDelay_month,AvgDepDelay_month) 
#Removing the common Month column
Avgdelayspermonth<-Avgdelayspermonth[,c(1,2,4)]
#Renaming the columns appropriately
colnames(Avgdelayspermonth)<-c('Month','AvgArrDelay','AvgDepDelay') 
#Fing the average of the average DepDelay and average ArrDelay
AvgofDelays <- (Avgdelayspermonth$AvgArrDelay+Avgdelayspermonth$AvgDepDelay)/2
#Binding them Avgdelaypermonth and AvgofDelays by columns
Avgdelayspermonth=cbind(Avgdelayspermonth,AvgofDelays)
#Viewing "Avgdelayspermonth"
show(Avgdelayspermonth)
```

#Plotting a lineplot to visualise
```{r}
#Using "ggplot2" library which is already imported above , to draw a line plot
library(ggplot2)
ggplot()+ geom_line(data=Avgdelayspermonth, aes(x=Month,y=AvgArrDelay, colour="AvgArrDelay"))+ geom_line(data=Avgdelayspermonth, aes(x=Month,y=AvgDepDelay, colour="AvgDepDelay")) + geom_line(data=Avgdelayspermonth, aes(x=Month,y=AvgofDelays, colour="AvgofDelays")) +
geom_point(data=Avgdelayspermonth, aes(x=Month,y=AvgArrDelay, colour="AvgArrDelay")) + geom_point(data=Avgdelayspermonth, aes(x=Month,y=AvgDepDelay, colour="AvgDepDelay")) + geom_point(data=Avgdelayspermonth, aes(x=Month,y=AvgofDelays, colour="AvgofDelays")) + xlab("Month")+ ylab("Average Delay/mins")+ scale_x_discrete(limits=c("Jan","Feb","March","April","May","June","July","August","Sep","Oct","Nov","Dec"))+
ggtitle("Line graph representing AvgDelays per month")
```
### September is the best Month to minimise delays

## Best Time Of Day to minimise delays

```{r}
#CRSDepTime column is converted into 8 different intervals and stored in a new column named"Timebucket" in the dataframe "data"
#Creating an empty column named "Timebucket"
data$Timebucket<- NA
#Creating values from 1 to 8 in "Timebucket" conditioning as follows on the "CRSDepTime" column
data[data$CRSDepTime>=0000 & data$CRSDepTime<0300,"Timebucket"] <- 1
data[data$CRSDepTime>=0300 & data$CRSDepTime<0600,"Timebucket"] <- 2
data[data$CRSDepTime>=0600 & data$CRSDepTime<0900,"Timebucket"] <- 3
data[data$CRSDepTime>=0900 & data$CRSDepTime<1200,"Timebucket"] <- 4
data[data$CRSDepTime>=1200 & data$CRSDepTime<1500,"Timebucket"] <- 5
data[data$CRSDepTime>=1500 & data$CRSDepTime<1800,"Timebucket"] <- 6
data[data$CRSDepTime>=1800 & data$CRSDepTime<2100,"Timebucket"] <- 7
data[data$CRSDepTime>=2100 & data$CRSDepTime<=2400,"Timebucket"] <- 8
```

##Plotting side by side boxplots to visualise
```{r}
#Creating a Totaldelay column in "data" dataframe
data$TotalDelay<- NA
data$TotalDelay=data$ArrDelay + data$DepDelay
```

```{r}
#Converting "Timebucket" into a factor to make it categorical
data$Timebucket<-as.factor(data$Timebucket)
```

#Using ggplot2 to do the box plot
```{r}
ggplot(data=data, aes(x=TotalDelay, y=Timebucket)) +
  geom_boxplot(fill="red") +xlim(c(-65,75))+
  ggtitle('Horizontal Boxplots of TotalDelay by Timeofday ')
```
### The time bucket "2" which is equal to [03:00,06:00) is the best time of day to minimise delays